---
title: "Coursera Proj 1"
output: html_document
---

For this assignment, the first step is to load the data file “activity.csv” by read.csv.
Next, a histogram of the daily total number of steps taken is generated, showing the distribution of these totals.

``` {r,echo=TRUE,warning=FALSE}
library(plyr)
df<-read.csv("activity.csv")
df1<-ddply(df,.(date),function(x) sum(x$steps))
hist(df1$V1,xlab = 'Steps')
```


Then we calculate the mean and the median of the number of steps taken.



```{r ,echo=TRUE,warning=FALSE}
m<-mean(df1$V1,na.rm = TRUE)
n<-median(df1$V1,na.rm = TRUE)
```

The mean number of steps per day is `r m` and the median of the number of steps per day is `r n` .


To exam the average daily activity pattern, we create a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r ,echo=TRUE,warning=FALSE}
df2<-tapply(df$steps ,df$interval , function(x) mean(x,na.rm = TRUE),simplify = TRUE)
df2<-data.frame(interval=names(df2),steps=df2)
with(df2,plot(interval,steps,type="l"))
```


Next is to check which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps:

```{r}
a<-which.max(df2$steps)
df2[a,]$interval
```

First, we calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs):

```{r}
length(which(! complete.cases(df)))

```


So the original data set has 2304 rows with missing data.

We use a simple strategy for filling in all of the missing values in the dataset. If a 5-minute interval has missing value, we use the mean for that 5-minute interval.

We create a new data frame df4 that is equal to the original dataset but with the missing data filled in (using mean for that interval for imputation):

```{r}
a<-complete.cases(df)
df3<-df[a,]
df4<-df
a<-is.na(df4$steps)
rval<-tapply(df3$steps ,df3$interval , mean)
df4$steps[a]<-rval[as.character(df4$interval[a])]
head(df4)
length(which(! complete.cases(df4)))
```


```{r,warning=FALSE}
df5<-ddply(df4,.(date),function(x) sum(x$steps))
hist(df5$V1,xlab = 'Steps')
m<-mean(df5$V1)
n<-median(df5$V1)
```

The new mean number of steps per day is `r m` and the median of the number of steps per day is `r n` .




Now,First we create a new factor variable “wk” in the dataset with two levels - “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
Next we make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
is_weekday <- function(d) {
    wd <- weekdays(d)
    ifelse (wd == "Saturday" | wd == "Sunday", "weekend", "weekday")
}

wx <- sapply(as.Date(df4$date), is_weekday)
df4$wk <- as.factor(wx)
head(df4)
wk_df <- aggregate(steps ~ wk+interval, data=df4, FUN=mean)

library(lattice)
xyplot(steps ~ interval | factor(wk),
       layout = c(1, 2),
       xlab="Interval",
       ylab="Number of steps",
       type="l",
       lty=1,
       data=wk_df)
```










